---
title: "SEM output"
author: "D van Denderen"
date: "8 March 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SEM model

SEM model is developed to analyse regional variation in demersal fish biomass. For more details on the demersal fish community data see <https://github.com/Dvandenderen/DemFish_trawl>.

SEM model is run for:

- Two spatial units
  * Ecoregions (n=20)
  * Management subdivisions (n=41)
  
- Three time periods
  * 1990-1995
  * 2000-2005
  * 2010-2015
  
Results are obtained with Rpackage Lavaan. The initial model has too many predictor variables for the number of observations. The initial model is therefore not evaluated for overall model fit - just used to analyse the pathways. The final model excludes depth and detrital flux. Both versions are used in the result section. Overall model fit is estimated using a Chi-square test (Lavaan) and Fisher's C test (piecewiseSEM). 

```{r,echo= T,  warning = F}
library(sp)
library(lavaan)
library(piecewiseSEM)
library(nlme)
```

SEM model causal pathway is primarily based on predictor variables used in the modelling of demersal fish in a global food web setup.

- Temperature is expected to decrease biomass (higher turnover, higher metabolic costs)
- Fishing exploitation is expected to decrease biomass
- More energy is lost in longer fish food webs (mean number of trophic steps)
- higher benthic and zoopl production increase fish biomass

```{r, warning = F, echo= T}
# load workspace (spatial analysis.R)
load (file = "C:/Users/danie/Documents/Online for git/DemFish_trawl/figures/my_work_space_up.RData")

# re-scale parameters
semparam <- c("NPP","lz_prod","biomass","tlw","SST_time","ER_log")

## hypothesized model lavaan
modT2='biomass ~ tlw + ER_log + SST_time + lz_prod
       tlw ~ NPP + ER_log
       lz_prod ~ NPP +  SST_time  
       NPP ~ SST_time'

## minimum nb of grid cells per region
id_min <- 6

```

**Subdivision, 1990-2015**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 1990,t_end = 2015,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + lz_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat
  
# re-scale parameters for Lavaan SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt1 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Subdivision, 1990-1995**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 1990,t_end = 1995,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + lz_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat
  
# re-scale parameters for Lavaan SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt2 <- summary(f.modT1.est,rsq=T,stand=T)

```


**Subdivision, 2000-2005**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2000,t_end = 2005,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + lz_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt3 <-summary(f.modT1.est,rsq=T,stand=T)

```

**Subdivision, 2010-2015**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2010,t_end = 2015,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + lz_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt4 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Ecoregion, 1990-2015**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 1990,t_end = 2015,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + lz_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat
  
# re-scale parameters for Lavaan SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt5 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Ecoregion, 1990-1995**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 1990,t_end = 1995,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + lz_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt6 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Ecoregion, 2000-2005**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2000,t_end = 2005,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + lz_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt7 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Ecoregion, 2010-2015**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2010,t_end = 2015,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + lz_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt8 <- summary(f.modT1.est,rsq=T,stand=T)

SEM_out = rbind(tt1$PE,tt2$PE,tt3$PE,tt4$PE,tt5$PE,tt6$PE,tt7$PE,tt8$PE)
write.csv(SEM_out,file="C:/Users/danie/Documents/Online for git/DemFish_trawl/figures/SEM_out.csv")

```

