---
title: "SEM output 50-250 meter depth"
author: "D van Denderen"
date: "1/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SEM model

SEM model is developed to analyse regional variation in demersal fish biomass. For more details on the demersal fish community data see <https://github.com/Dvandenderen/DemFish_trawl>.

SEM model is run for:

- Two spatial units
  * Ecoregions (n=20)
  * Management subdivision (n=41)
  
- Three time periods
  * 1990-1995
  * 2000-2005
  * 2010-2015
  
Results are obtained with Rpackage Lavaan 
```{r,echo= T,  warning = F}
library(sp)
library(lavaan)
```

SEM model causal pathway is primarily based on predictor variables used in the modelling of demersal fish in a global food web setup.

- Temperature is expected to decrease biomass (higher turnover, higher metabolic costs)
- Fishing exploitation is expected to decrease biomass
- More energy is lost in longer fish food webs (mean number of trophic steps)
- higher benthic and zoopl production increase fish biomass
```{r, warning = F, echo= T}
# load workspace (spatial analysis.R)
load (file = "C:/Users/danie/Documents/Online for git/DemFish_trawl/figures/my_work_space.RData")

# create dataset by selecting depths 50 - 250 m
depth_grid$uni <- paste(depth_grid$uni_cell,depth_grid$year)
cpue_depth <- cbind(cpue_good,depth_grid[match(cpue_good$uni,depth_grid$uni),c(3)])
colnames(cpue_depth)[ncol(cpue_depth)] <- "depth"
cpue_depth <- subset(cpue_depth,cpue_depth$depth > 50 & cpue_depth$depth < 250)
cpue_good <- subset(cpue_good,cpue_good$uni %in% cpue_depth$uni)

# re-scale parameters
semparam <- c("NPP","lz_prod","ben_prod","biomass","tlw","depth","SST_time","ER_log")

## hypothesized model 
modT2='biomass ~ tlw + ER_log + SST_time + ben_prod + lz_prod
       tlw ~ NPP + ER_log
       lz_prod ~ NPP +  SST_time 
       ben_prod ~ NPP + SST_time + depth+lz_prod
       NPP ~ SST_time'

## minimum nb of grid cells per region
id_min <- 6

```

**Subdivision, 1990-1995**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 1990,t_end = 1995,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt1 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Subdivision, 2000-2005**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2000,t_end = 2005,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt2 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Subdivision, 2010-2015**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2010,t_end = 2015,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt3 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Ecoregion, 1990-1995**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 1990,t_end = 1995,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt4 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Ecoregion, 2000-2005**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2000,t_end = 2005,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt5 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Ecoregion, 2010-2015**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2010,t_end = 2015,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt6 <- summary(f.modT1.est,rsq=T,stand=T)

SEM_out = rbind(tt1$PE,tt2$PE,tt3$PE,tt4$PE,tt5$PE,tt6$PE)
write.csv(SEM_out,file="C:/Users/danie/Documents/Online for git/DemFish_trawl/figures/SEM_out50_250.csv")

```

