---
title: "SEM output"
author: "D van Denderen"
date: "8 March 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SEM initial model

SEM model is developed to analyse regional variation in demersal fish biomass. For more details on the demersal fish community data see <https://github.com/Dvandenderen/DemFish_trawl>.

SEM model is run for:

- Two spatial units
  * Ecoregions (n=20)
  * Management subdivisions (n=41)
  
- Three time periods
  * 1990-1995
  * 2000-2005
  * 2010-2015
  
The initial model has too many predictor variables for the number of observations. The initial model is therefore not evaluated for overall model fit - just used to analyse the pathways. The final model excludes depth and detrital flux. Both versions are used in the result section. Output file of the final model is "SEM final - knitted output". 

```{r,echo= T,  warning = F}
library(sp)
library(piecewiseSEM)
```

Initial SEM model causal pathway is primarily based on predictor variables used in the modelling of demersal fish in a global food web setup.

- Temperature is expected to decrease biomass (higher turnover, higher metabolic costs)
- Fishing exploitation is expected to decrease biomass
- More energy is lost in longer fish food webs (mean number of trophic steps)
- higher benthic and zoopl production increase fish biomass
```{r, warning = F, echo= T}
# load workspace (spatial analysis.R)
load (file = "C:/Users/danie/Documents/Online for git/DemFish_trawl/figures/my_work_space_up.RData")

## minimum nb of grid cells per region
id_min <- 6

```

**Subdivision, 1990-1995**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 1990,t_end = 1995,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + ben_prod + lz_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(ben_prod ~ NPP + SST_time + depth +lz_prod, data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

summary(model, .progressBar = F)

```


**Subdivision, 2000-2005**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2000,t_end = 2005,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + SST_time  + ER_log + lz_prod + ben_prod, data=datbio),
  lm(tlw ~ ER_log + NPP, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(ben_prod ~NPP + SST_time + depth + lz_prod, data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

summary(model, .progressBar = F)

```

**Subdivision, 2010-2015**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2010,t_end = 2015,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + SST_time  + ER_log + lz_prod + ben_prod, data=datbio),
  lm(tlw ~ ER_log + NPP, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(ben_prod ~NPP + SST_time + depth + lz_prod, data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

summary(model, .progressBar = F)

```

**Ecoregion, 1990-1995**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 1990,t_end = 1995,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + SST_time  + ER_log + lz_prod + ben_prod, data=datbio),
  lm(tlw ~ ER_log + NPP, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(ben_prod ~NPP + SST_time + depth + lz_prod, data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

summary(model, .progressBar = F)

```

**Ecoregion, 2000-2005**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2000,t_end = 2005,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + SST_time  + ER_log + lz_prod + ben_prod, data=datbio),
  lm(tlw ~ ER_log + NPP, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(ben_prod ~NPP + SST_time + depth + lz_prod, data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

summary(model, .progressBar = F)

```

**Ecoregion, 2010-2015**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2010,t_end = 2015,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + SST_time  + ER_log + lz_prod + ben_prod, data=datbio),
  lm(tlw ~ ER_log + NPP, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(ben_prod ~NPP + SST_time + depth + lz_prod, data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

summary(model, .progressBar = F)
```

Results show that the benthic flux to biomass is 6/6 times non-significant. Re-doing the analysis for the entire time series generally supports the above findings (trophic level is now not very important). Final model is simplified and  depth and detrital bottom flux are excluded.  

**Subdivision, 1990-2015**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 1990,t_end = 2015,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + ben_prod + lz_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(ben_prod ~ NPP + SST_time + depth +lz_prod, data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

summary(model, .progressBar = F)

```

**Ecoregion, 1990-2015**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 1990,t_end = 2015,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + SST_time  + ER_log + lz_prod + ben_prod, data=datbio),
  lm(tlw ~ ER_log + NPP, data=datbio),
  lm(lz_prod ~ NPP +  SST_time , data=datbio),
  lm(ben_prod ~NPP + SST_time + depth + lz_prod, data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

summary(model, .progressBar = F)

```
