---
title: "SEM output"
author: "D van Denderen"
date: "August 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SEM model grid cell sensitivity

SEM model is developed to analyse regional variation in demersal fish biomass. For more details on the demersal fish community data see <https://github.com/Dvandenderen/DemFish_trawl>.

SEM model is run for:

- Two spatial scales
  * Ecoregions (n=21)
  * Management subdivisions (n=45)
  
- Three time periods
  * 1990-1995
  * 2000-2005
  * 2010-2015
  
Results are obtained with Rpackage Lavaan. Not all grid cells had data in all time periods. In the main analysis, we removed all ecoregions/subdivisions with fewer than 6 grid cells with observations. Here, we include all data.

```{r,echo= T,  warning = F}
library(sp)
library(lavaan)
library(piecewiseSEM)
library(nlme)
```

SEM model causal pathway is primarily based on predictor variables used in the modelling of demersal fish in a global food web setup.

- Temperature is expected to decrease biomass (higher turnover, higher metabolic costs)
- Fishing exploitation is expected to decrease biomass
- More energy is lost in longer fish food webs (mean number of trophic steps)
- zoopl production increases fish biomass

```{r, warning = F, echo= T}
# load workspace (spatial analysis.R)
load (file = "C:/Users/danie/Documents/Online for git/DemFish_trawl/figures/my_work_space.RData")

# re-scale parameters
semparam <- c("NPP","z_prod","biomass","tlw","SST_time","ER_log")

## hypothesized model lavaan
modT2='biomass ~ tlw + ER_log + SST_time + z_prod
       tlw ~ NPP + ER_log
       z_prod ~ NPP +  SST_time  
       NPP ~ SST_time
       SST_time ~~ 0*ER_log'

## minimum nb of grid cells per region
id_min <- 1

```

**Subdivision, 1990-2015**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 1990,t_end = 2015,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# get total zooplankton production
datbio$z_prod <- datbio$lz_prod + datbio$mz_prod

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + z_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(z_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat
  
# re-scale parameters for Lavaan SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt1 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Subdivision, 1990-1995**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 1990,t_end = 1995,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# get total zooplankton production
datbio$z_prod <- datbio$lz_prod + datbio$mz_prod

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + z_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(z_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat
  
# re-scale parameters for Lavaan SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt2 <- summary(f.modT1.est,rsq=T,stand=T)

```


**Subdivision, 2000-2005**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2000,t_end = 2005,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# get total zooplankton production
datbio$z_prod <- datbio$lz_prod + datbio$mz_prod

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + z_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(z_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt3 <-summary(f.modT1.est,rsq=T,stand=T)

```

**Subdivision, 2010-2015**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2010,t_end = 2015,spatialunit = "subdivision")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# get total zooplankton production
datbio$z_prod <- datbio$lz_prod + datbio$mz_prod

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + z_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(z_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt4 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Ecoregion, 1990-2015**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 1990,t_end = 2015,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# get total zooplankton production
datbio$z_prod <- datbio$lz_prod + datbio$mz_prod

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + z_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(z_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat
  
# re-scale parameters for Lavaan SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt5 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Ecoregion, 1990-1995**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 1990,t_end = 1995,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# get total zooplankton production
datbio$z_prod <- datbio$lz_prod + datbio$mz_prod

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + z_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(z_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt6 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Ecoregion, 2000-2005**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2000,t_end = 2005,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# get total zooplankton production
datbio$z_prod <- datbio$lz_prod + datbio$mz_prod

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + z_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(z_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt7 <- summary(f.modT1.est,rsq=T,stand=T)

```

**Ecoregion, 2010-2015**

```{r }
# source data, year start, year end, spatialunit = "ECO_REG" or "subdivision"
datbio <- get_bio(cpue_good,t_start = 2010,t_end = 2015,spatialunit = "ECO_REG")  
datbio <- subset(datbio,datbio$countgrid >= id_min) # exclude regions with few grid cells

# get exploitation rate
datbio$ER_log <- log10(datbio$Catch_sqkm/datbio$biomass)

# get total zooplankton production
datbio$z_prod <- datbio$lz_prod + datbio$mz_prod

# piecewise SEM
model <- psem(
  lm(biomass ~ tlw + ER_log + SST_time + z_prod, data=datbio),
  lm(tlw ~ NPP + ER_log, data=datbio),
  lm(z_prod ~ NPP +  SST_time , data=datbio),
  lm(NPP ~ SST_time , data=datbio),
  datbio)

results <- summary(model, .progressBar = F)
results$Cstat

# re-scale parameters for SEM
datbio[,semparam] <- scale(datbio[,semparam])

# SEM 
f.modT1.est <- sem(modT2,fixed.x=F,data=datbio)
tt8 <- summary(f.modT1.est,rsq=T,stand=T)

SEM_out = rbind(tt1$PE,tt2$PE,tt3$PE,tt4$PE,tt5$PE,tt6$PE,tt7$PE,tt8$PE)
write.csv(SEM_out,file="C:/Users/danie/Documents/Online for git/DemFish_trawl/figures/SEM_out_grid1.csv")

```

